---
title: "Overview"
author: "Janko Thyson"
date: "Wednesday, February 04, 2015"
output: html_document
---

```{r, echo=FALSE}
library("reposr")
```

## Typical workflow 

### Specifiy root directory 

We will use the following directory as our local CRAN-like package repoitory:

```{r}
root <- file.path(tempdir(), "cran")
```

### Create instance

```{r}
repo <- PackageRepository$new(root)
```

### Create/ensure repository

```{r}
repo$ensure()
```

## Verify existence 

```{r}
repo$exists()
```

### Register as R option

```{r}
(getOption("repos"))
repo$register()
(getOption("repos"))
```

### Browse content 

```{r}
repo$browse()
```

### Built into 

```{r}
repo$buildInto()
```

### Investigate

Show content based on index file:

```{r}
repo$show()
repo$show(type = "source")
repo$show(type = "mac.binary")
repo$show(type = "win.binary")
```

Check if specific package(s) exist(s):

```{r}
repo$has()
repo$has(type = "source", atomic = FALSE)
repo$has(c("devtools", "dplyr"))
```

### Maintain

Remove outdated packages and refresh. Outdated packages are moved to an 
special repository archive `repo$root_atomic`. 

Each outdated package build will be integrated into its own "one-package-only"  repository: `file.path(repo$root_atomic, "<pkg_name>", "<pgk_version>")`

```{r}
repo$clean()
```

Remove packages:

```{r}
repo$remove()
repo$remove("<pkg_name>")
```

Reset entire repository:

```{r}
repo$reset()
```

### Delete

```{r}
repo$exists()
repo$delete()
repo$exists()
```
