---
title: "Reference"
author: "Janko Thyson"
date: "Wednesday, February 04, 2015"
output: html_document
---

```{r, echo=FALSE}
library("repor")
```


## Specifiy root directory 

We will use the following directory as our local CRAN-like package repoitory:

```{r}
root <- file.path(tempdir(), "cran")
```

## Create instance

```{r}
repo <- PackageRepository$new(root)
```

## Object structure

As the most important feature that characterizes a (local) package repoitory is (arguably) its location, the print method for class `PackageRepository` only shows the value of field `root` instead of what is returned by the default method for `R6` objects (`print.R6`). This value corresponds to the package repoitory's root directory:

```{r}
repo
```

In order to get an idea about the actual object structure, you can call the print method explicitly and use `full = TRUE`:

```{r}
print(repo, full = TRUE)
```

The most important subdirectories that are actively linked to `$root` (and the currently used minor version of R) are:

```{r}
repo$win.binary
repo$mac.binary
repo$source
```


## Typical workflow

### As URL (`$asUrl`)

Often times, you need your local repoitory to be in an [URL format](http://en.wikipedia.org/wiki/Uniform_resource_locator) with 
[URL scheme](http://en.wikipedia.org/wiki/URI_scheme) `file` (also see `?url`)

```{r}
repo$asUrl()
```

Other [URL scheme](http://en.wikipedia.org/wiki/URI_scheme) are also implement, but not really useful as this package focuses on repoitories on a *local* file system (as opposed to a true web-based ressource for which schemes `http` and `ftp` are normally used).

```{r}
repo$asUrl("http")
repo$asUrl("ftp")
```

### Create/ensure (`$ensure()`)

Create/ensure a local package repoitory that complies with the structural conventions of [CRAN](http://cran.r-project.org/):

```{r}
repo$ensure()
repo$ensure(plain = TRUE)
## --> unnamed
```

### Delete (`$delete()`)

Delete root directory including all subdirectories. 

As for all methods involving critical operations (in the sense that you wouldn't want to perform them accidentally), this method takes an `ask` argument that prompts the user for confirmation:

```{r}
repo$delete()
```

Whenever user is prompted for input, the valid inputs are given at the end of the line (`yes` [or `y`], `no` [or `n`] and `quit` [or `q`] in this and most cases)

You can suppress this by setting `ask = FALSE`

```{r}
repo$delete(ask = FALSE)
repo$delete(ask = FALSE, plain = TRUE)
## --> unnamed
```

What happens when you try to delete a repoitory that does not exist depends on the settings for `strict`. It defaults to `0` in which case no additional message (or condition is issued):

```{r}
repo$delete()
```

Here is what happens when `strict` is `> 0`:

```{r}
repo$delete(strict = 1)
repo$delete(strict = 2)
repo$delete(strict = 3)
```


### Check existence (`$exists()`)

```{r}
repo$exists()
repo$exists(plain = TRUE)
repo$ensure()
repo$exists()
repo$delete(ask = FALSE)
repo$exists()
```

### Build into (`$buildInto()`)

You can seamlessly build your package directly into a local package repoitory using functions of package [`devtools`](https://github.com/hadley/devtools).

First, we ensure the repoitory's existence:

```{r}
repo$ensure()
```

Next, we use the `browse` method to display the current repoitory content via the OS's default files system browser: 

```{r}
repo$browse(getOption("pkgType"))
```

Before we build our package, let's see the current content of the repoitory's index file:

```{r}
repo$show()
```

As we see, the index is still empty at this point.

Now we build our package into the local repoitory. Note that this step assumes that your working directory points to a valid R package project. By default, the *source* version of the package is always along with the binary version corresponding to `getOption("pkgType")` (if your OS is either [Windows](http://windows.microsoft.com/de-de/windows-8/why-windows) or [Mac OS](https://www.apple.com/de/osx/))

```{r}
repo$buildInto()
```

Note that the repoitory directory's content has changed: the repoitory now contains the package build(s) and the respective index file(s) (`PACKAGES`) has/have changed.

The index has changed as well:

```{r}
repo$show()
```


### Check existence of specific package(s) (`$has()`)

This step assumes that you built a package into the repoitory via `$buildInto()`.

When your working directory points to a valid package project, then the method automatically identifies the package name based on the project's `DESCRIPTION` file when being called without an explicit value for `pkg`):

```{r}
repo$has()
```

When the name can't be retrieved, an error is thrown:

```{r}
wd <- setwd("man") 
## --> unexpected working directory
repo$has()

## Clean up //
setwd(wd)
```

Specify explicit values for `pkg`:

```{r}
my_pkg <- "repor" 
## Please change to your package's name

repo$has(my_pkg)

## To include the information about the package type,
## use `atomic = FALSE`:
repo$has(my_pkg, atomic = FALSE)

repo$has(my_pkg, type = "source", atomic = FALSE)
repo$has(my_pkg, type = "mac.binary", atomic = FALSE)

repo$has("devtools")
repo$has(c(my_pkg, "devtools"))
```

## Second-order methods for users 

### Are there any packages at all (`$hasAny()`)

Check if a repoitory contains *any* packages at all

```{r}
repo$hasAny()
repo$hasAny(atomic = TRUE)
```

### Show content (`$show()`)

Show the content based on a repoitory's index file.

This step assumes that you built a package into the repoitory via `$buildInto()`.

```{r}
repo$ensure()
repo$show()
repo$show(type = "win.binary")
repo$show(type = "mac.binary")
repo$show(type = "source")
```

### Refresh (`$refresh()`)

Refresh a repoitory. This implies updating the index files according to the latest available package builds.

```{r}
repo$ensure()
repo$refresh()
```

### Remove packages (`$remove`)

This step assumes that you built a package into the repoitory via `$buildInto()`.

Remove package associated to the current package project (across all package type subdirectories):

```{r}
repo$browse(getOption("pkgType"))
## --> note the different status before and after removal

repo$remove()
```


With explicit value for `pkg` (multiple values are also possible):

```{r}
my_pkg <- "repor"
## Please change to your package's name

repo$remove(my_pkg)
repo$remove(my_pkg, ask = FALSE)
## --> already removed
```

Remove packages selectively (types/subdirectories). This requires a re-build of your package.

```{r}
my_pkg <- "repor"
## Please change to your package's name

repo$buildInto()
repo$browse(path = "source")

repo$remove(my_pkg, type = "source", ask = FALSE)
## --> only the source build is removed

repo$browse(path = getOption("pkgType"))
repo$remove(my_pkg, type = getOption("pkgType"), ask = FALSE)
## --> only the default package type is removed
```

### Reset (`$reset()`)

Resets a repoitory to the initial state (corresponds to the result of initially calling `$ensure()` or `$ensure(overwrite = TRUE)`).

```{r}
repo$buildInto()
repo$browse(path = "source")
repo$reset()
```

### Clean (`$clean()`)

Cleans a repoitory in the sense that old package versions are either deleted or archived. By default, a repoitory is refreshed before the actual cleaning takes place (can be controlled via `refresh = TRUE/FALSE`).

```{r}
repo$clean()
```

Old versions are archived to directory `<repo-name>_archive` below the parent directory a repoitory's root directory

In our case, this would be the following directory
```{r}
paste0(repo$root, "_archive")
```

#### Note

You can delete repoitory archives by using the `archive` parameter:

```{r}
repo$delete(archive = TRUE)
```

